HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGET BRAIN ALGORITHM

### 2012-10-27 paper.tex
Updated a little of the methods, and added color/hyperref packages.

\documentclass{article}
\usepackage{spconf,amsmath,graphicx}

\newcommand{\todo}[1]{
    \addcontentsline{tdo}{todo}{\protect{#1}}
    \textcolor{red}{#1}
}

\title{HIPPOCAMPAL SEGMENTATION}
\name{  Jon Pipitone$^{\star}$ Jason P. Lerch$^{\dagger}$ Miriam Friedel$^{\dagger}$ Aristotle Voineskos$^{\star}$
         Mallar Chakravarty$^{\star}$ }

@@

\begin{document}
\SweaveOpts{concordance=TRUE}
\maketitle

@@

We use the {\it cross-correlation weighted voting} method to fuse the
candidate labels of each unlabeled image.  In this method, a voxel-wise
majority vote is carried out the labels from the top $n$ ranked template
library images.  The scores are given by the normalized cross-correlation
between the unlabeled image and each template library image, after the
unlabeled image is linearly registered to the template image.  The
cross-correlation score itself is calculated over a region of interest
(ROI) defined on each template.  A customized ROI is defined by first
linearly registering each unlabeled subject to all possible templates in
the template library, transforming all template hippocampus labels,
defining the possible extent of all hippocampal labels, and finally
dilating this extent by three voxels.
\todo{jp: it isn't clear that only the top ranked images are being used.
perhaps some rationale for CC voting would help here. i.e. discuss improvement
in overall registration quality --  throwing out atlases liable to introduce
unwanted variance }

@@

For evaluation purposes we used the publicly available IBSR dataset.  This
dataset consists of T1 weighted MR image volumes from 18 subjects (4 females,
14 males) with ages between 7 and 71 years. Image dimensions for all MR volumes
are normalized to  $256  \times 256 \times 128$ voxels, with the voxel size
ranging from $0.8 \times 0.8 \times 1.5 mm$ to $1.0 \times 1.0 \times 1.5 mm$.
The images come 'positionally normalized' into the Talairach orientation
(rotation only), and processed by the CMA 'autoseg' biasfield correction
routines. The MR brain data sets and their manual segmentations are publicly
available and were provided by the Center for Morphometric Analysis at
Massachusetts General Hospital and are available at
\url{http://www.cma.mgh.harvard.edu/ibsr/}.

@@

\begin{figure}
\begin{minipage}[b]{1.0\linewidth}
  \centering
  %\includegraphics[width=\textwidth]{results}
<<fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE, >>=
library(ggplot2)
results = read.csv(file='data/results.csv')
xcorr_vote= subset(results, method=='xcorr' | method == 'multiatlas_xcorr_vote' )
qplot(x=top_n, y=k_hippocampus, data=xcorr_vote, colour=as.factor(num_atlases), geom=c('point'), alpha=I(0)) +
  geom_smooth(line=1.5, formula=y~poly(x,2), method='lm') + 
  geom_hline(aes(yintercept = mean(subset(results, method=='naive')$k_hippocampus))) + 
  theme(legend.position="none") + 
  ylab('Kappa') + 
  xlab('Number of Atlases') +
  annotate("text", 16.0, 0.712, label="Naive", size=4, hjust=0)+
  annotate("text", 15.3, 0.754, label="3 atlases", size=4, hjust=0)+
  annotate("text", 14.4, 0.766, label="4 atlases", size=4, hjust=0)+
  annotate("text", 13.5, 0.770, label="5 atlases", size=4, hjust=0)+
  annotate("text", 12.5, 0.773, label="6 atlases", size=4, hjust=0)+
  annotate("text", 11.5, 0.776, label="7 atlases", size=4, hjust=0)+
  annotate("text", 10.3, 0.779, label="8 atlases", size=4, hjust=0)+
  annotate("text", 14.6, 0.793, label="Multi-atlas", size=4, hjust=0)+
  coord_cartesian(ylim=c(0.7,0.8), xlim=c(3,18)) + 
  scale_y_continuous(breaks=seq(0,1,0.01))

@  
\end{minipage}
\caption{{\em Performance of MAGeT Brain across all validations compared to
mean naive segmentation, and mean multi-atlas with cross-correlation voting.}
Multi-atlas and naive segmentations procedures provide upper and lower
bounds (respectively) for accuracy.  Note that improved segmentation
accuracy is achieved through increasing the number of input atlases.}
\label{results}
\end{figure}

>>>>

\documentclass{article}
\usepackage{spconf,amsmath,graphicx,color,hyperref}

\newcommand{\todo}[1]{
    \textcolor{red}{#1}
}

\title{HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGET
BRAIN ALGORITHM} 
\name{  Jon Pipitone$^{\star}$ Jason P. Lerch$^{\dagger}$ Miriam Friedel$^{\dagger}$ Aristotle Voineskos$^{\star}$
         Mallar Chakravarty$^{\star}$ }

@@

\usepackage{Sweave}
\begin{document}
\input{paper-concordance}
\maketitle

@@

We use the {\it cross-correlation weighted voting} method to both improve the
quality of the label fusion step and reduce the number of registrations
required \todo{ref}.  In this method, each template library image is ranked 
in similarity to each unlabelled image, and only the top $n$ template library
images labels are used for the voxel-wise majority vote fusion step.
Similarity scores are computed by the normalized cross-correlation of image
intensities between an unlabeled image and each template library image, after
the unlabeled image has undergone linearly registration to the template image.
The cross-correlation score is calculated over a region of interest (ROI) on
each template representing an generous approximation of hippocampal structure
and neighbourhood.  The ROI is defined as the extent of all atlas labels after
linear registration to the template, dilated by three voxels.

@@

For evaluation purposes we used the publicly available IBSR dataset.  This
dataset consists of T1 weighted MR image volumes from 18 subjects (4 females,
14 males) with ages between 7 and 71 years. Image dimensions for all MR volumes
are normalized to  $256  \times 256 \times 128$ voxels, with the voxel size
ranging from $0.8 \times 0.8 \times 1.5 mm$ to $1.0 \times 1.0 \times 1.5 mm$.
The images come 'positionally normalized' into the Talairach orientation
(rotation only), and processed by the CMA 'autoseg' biasfield correction
routines. The MR brain data sets and their manual segmentations are publicly
available and were provided by the Center for Morphometric Analysis at
Massachusetts General Hospital and are available at
\url{''http://www.cma.mgh.harvard.edu/ibsr/''}.

@@

\begin{figure}
\begin{minipage}[b]{1.0\linewidth}
  \centering
  %\includegraphics[width=\textwidth]{results}
\includegraphics{paper-001}
\end{minipage}
\caption{{\em Performance of MAGeT Brain across all validations compared to
mean naive segmentation, and mean multi-atlas with cross-correlation voting.}
Multi-atlas and naive segmentations procedures provide upper and lower
bounds (respectively) for accuracy.  Note that improved segmentation
accuracy is achieved through increasing the number of input atlases.}
\label{results}
\end{figure}
